---
title: "Time Series Project - Exploratory Analysis"
author: "Jainam Mehta"
date: "16 May 2020"
output: rmarkdown::github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Import necessary libraries
```{r message=FALSE, warning=FALSE}
library(tseries)
library(forecast)
```

## Data Prep

Read in the data
```{r message=FALSE, warning=FALSE}
library(Quandl)
Quandl.api_key("UqYbx-p5DwWzyT8AH2c7")
marketpriceUSD<-Quandl("BCHAIN/MKPRU")
```

Reverse the series to get prices in chronological order
```{r}
chron <- rev(marketpriceUSD$Value)
c(start = tail(marketpriceUSD$Date, 1), end = head(marketpriceUSD$Date, 1))
```

Covert to time series object
```{r}
BTC <- ts(chron, start = c(2009, 1, 3), frequency = 365)

# focus on 2016 onwards
train <- window(BTC, start = 2016, end  = 2020)
test <- window(BTC, start = 2020)
```

Plot the data and run tests for stationarity
```{r}
plot(train)

acf(train); pacf(train)

adf.test(train)
kpss.test(train)
```
The plot, along with both ADF and KPSS tests suggest that the series is definitely non-stationary.

Take first differences to see if the series can be made stationary
```{r}
train.diff <- diff(train, lag = 1)
plot(train.diff)
adf.test(train.diff)
kpss.test(train.diff)
```
Taking a difference with lag = 1 seems to make the series stationary.


## Model Fitting

Fit an ARIMA model using auto.arima() function
```{r}
(fit <- auto.arima(train, allowdrift = T))
forecast <- forecast(fit, h=138)
plot(forecast)
```

Check residuals
```{r}
residuals <- test - forecast$mean
mean(residuals)
checkresiduals(residuals)
```

DirRec method of forecasting
```{r}
train.append <- train
dirrec.forecast <- c()
dirrec.model <- list()

for (i in 1:30) {
  model <- auto.arima(train.append)
  dirrec.model[[i]] <- model
  f <- forecast(model, h = 1)$mean
  dirrec.forecast <- c(dirrec.forecast, f)
  train.append <- ts(c(train.append, f), start = c(1975, 1), frequency = 12)
}
```

Compare the Recursive and DirRec forecasts
```{r}
plot(1:30, test[1:30], type = 'l', lwd = 2,
     main = 'Bitcoin price in Jan 2020',
     xlab = 'January', ylab = 'Price')
lines(1:30, forecast$mean[1:30], type = 'l', lwd = 2, col = 'purple')
lines(1:30, dirrec.forecast, type = 'l', lwd = 2, col = 'orange')
legend("topright", legend=c("Actual", "Rec", "DirRec"),
       col=c("black", "purple", "orange"), lty=1, cex=0.8)

#forecast$mean[1:30]
#dirrec.forecast
```









