---
title: "Time Series Project - Multivariate Analysis"
author: "Jainam Mehta"
date: "02 June 2020"
output: rmarkdown::github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Import necessary libraries
```{r message=FALSE, warning=FALSE}
library(tseries)
library(forecast)
library(tidyverse)
library(MTS)
library(imputeTS)
library(TSA)
library(lubridate)
library(vars)
```

## Data Prep

Download various datasets from Quandl

* BTC Market price (https://www.quandl.com/data/BCHAIN/MKPRU-Bitcoin-Market-Price-USD)
* ETH to USD rate (https://www.quandl.com/data/GDAX/ETH_USD-ETH-USD-Exchange-Rate)
* LTC to USD rate (https://www.quandl.com/data/GDAX/LTC_USD-LTC-USD-Exchange-Rate)
* ETH to BTC rate (https://www.quandl.com/data/GDAX/ETH_BTC-ETH-BTC-Exchange-Rate)
* LTC to BTC rate (https://www.quandl.com/data/GDAX/LTC_BTC-LTC-BTC-Exchange-Rate)
* Bitcoin Miners Revenue (https://www.quandl.com/data/BCHAIN/MIREV-Bitcoin-Miners-Revenue)
* Number of Transactions (https://www.quandl.com/data/BCHAIN/NTRAN-Bitcoin-Number-of-Transactions)

```{r message=FALSE, warning=FALSE}
library(Quandl)
Quandl.api_key("UqYbx-p5DwWzyT8AH2c7")

# Select the relevant series
BTC_USD_raw<-Quandl("BCHAIN/MKPRU")
ETH_USD_raw<-Quandl("GDAX/ETH_USD")
LTC_USD_raw<-Quandl("GDAX/LTC_USD")
ETH_BTC_raw<-Quandl("GDAX/ETH_BTC")
LTC_BTC_raw<-Quandl("GDAX/LTC_BTC")

# Select non-price predictors
MIREV_raw<-Quandl("BCHAIN/MIREV")
NTRAN_raw<-Quandl("BCHAIN/NTRAN")
```


Create a function to convert the data into proper time series objects. 
This function reverses the time series so that it is chrononlogical and then assigns it to a time series object with the correct start date.
```{r}
create_ts <- function(rawdata) {
  # use opening price, if Value column doesn't exist
  if (!"Value" %in% colnames(rawdata)) {rawdata$Value <- rawdata$Open} 
  
  rawdata$Date <- as.Date(rawdata$Date)
  # fill missing dates with NA.
  rawdata <- complete(rawdata, Date = seq(min(Date), max(Date), by = "day")) 
  #year <- as.numeric(format(rawdata$Date,'%Y'))[1]
  #month <- as.numeric(format(rawdata$Date,'%m'))[1]
  #day <- as.numeric(format(rawdata$Date,'%d'))[1]
  
  return(ts(rawdata$Value, decimal_date(as.Date(rawdata$Date[1])), frequency = 365))
}
```


Create time series objects for all data
```{r}
BTC_USD <- create_ts(BTC_USD_raw)
plot(BTC_USD)

ETH_USD <- create_ts(ETH_USD_raw)
plot(ETH_USD)

LTC_USD <- create_ts(LTC_USD_raw)
plot(LTC_USD)

ETH_BTC <- create_ts(ETH_BTC_raw)
plot(ETH_BTC)

LTC_BTC <- create_ts(LTC_BTC_raw)
plot(LTC_BTC)

MIREV <- create_ts(MIREV_raw)
plot(MIREV)

NTRAN <- create_ts(NTRAN_raw)
plot(NTRAN)

```

Function to create imputed series
```{r}
series_names <- list(BTC_USD, ETH_USD, LTC_USD, ETH_BTC, LTC_BTC, MIREV, NTRAN)
train <- matrix()
test <- matrix()

for (i in series_names) {
  imputed <- na_interpolation(i, option = 'linear')
  plotNA.imputations(i, imputed)
  tr <- window(imputed, start = 2018, end = 2019)
  plot(tr)
  te <- window(imputed, start = 2020, end = c(2020, 90))
  plot(te)
  test <- cbind(test, te)
}
```

Manually created imputed series for ETH_USD and LTC_USD
```{r}
ETH_USD.imp <- na_interpolation(ETH_USD, option = 'spline')
plotNA.imputations(ETH_USD, ETH_USD.imp)

LTC_USD.imp <- na_interpolation(LTC_USD, option = 'spline')
plotNA.imputations(LTC_USD, LTC_USD.imp)
```

Take log of the data
```{r}
BTC_USD <- log(BTC_USD)
ETH_USD <- log(ETH_USD)
LTC_USD <- log(LTC_USD)
```


Create train and test splits for each
```{r}
BTC_USD.train <- window(BTC_USD, start = 2018, end = 2020)
MIREV.train <- window(MIREV, start = 2018, end = 2020)
NTRAN.train <- window(NTRAN, start = 2018, end = 2020)
ETH_USD.train <- window(ETH_USD.imp, start = 2018, end = 2020)
LTC_USD.train <- window(LTC_USD.imp, start = 2018, end = 2020)

BTC_USD.test <- window(BTC_USD, start = 2020, end = c(2020, 90))
MIREV.test <- window(MIREV, start = 2020, end = c(2020, 90))
NTRAN.test <- window(NTRAN, start = 2020, end = c(2020, 90))
ETH_USD.test <- window(ETH_USD.imp, start = 2020, end = c(2020, 90))
LTC_USD.test <- window(LTC_USD.imp, start = 2020, end = c(2020, 90))

plot(BTC_USD.train, main = 'Train')
plot(BTC_USD.test, main = 'Test')
```



# Multivariate Analysis

Select the 3 series of the market prices of BTC, ETH and LTC.
```{r}
cryptos <- na.omit(cbind(BTC = BTC_USD.train, ETH = ETH_USD.train, LTC = LTC_USD.train))
head(cryptos)
plot(cryptos, main = 'Crypto Market Price in USD')

cor(cryptos)
```
Not that the series are all quite correlated as expected.


Run tests for stationarity.
```{r}
apply(cryptos, 2, adf.test)
apply(cryptos, 2, kpss.test)
```

Since the series are all non-stationary, we take first differences
```{r}
cryptos.stnry <- diffM(cryptos)

# Run stationarity tests to check now
apply(cryptos.stnry, 2, adf.test)
#apply(cryptos.stnry, 2, kpss.test)
```

Plot the differenced series
```{r}
plot.ts(cryptos.stnry)
```

Identify the Lag orders.
We will use two different functions, from two different packages to identify the lag order for the VAR model. Both functions are quite similar to each other but differ in the output they produce. vars::VAR is a more powerful and convinient function to identify the correct lag order. 
```{r}
VARselect(cryptos.stnry, 
          type = "none", #type of deterministic regressors to include. We use none becasue the time series was made stationary using differencing above. 
          lag.max = 10) #highest lag order

```

Creating a VAR model with vars
```{r}
var.a <- vars::VAR(cryptos.stnry,
                   lag.max = 10, #highest lag order for lag length selection according to the choosen ic
                   ic = "AIC", #information criterion
                   type = "none") #type of deterministic regressors to include
summary(var.a)


?VAR
```

```{r}
var.a$p
```




Residual diagnostics
```{r}
#serial.test function takes the VAR model as the input.  
serial.test(var.a)
```

Run Granger test for causality. We want to check if ETH and LTC prices do indeed affect BTC or not
```{r}
# for causality function to give reliable results we need all the variables of the multivariate time series to be stationary. 
causality(var.a, cause = c("ETH", "LTC")) #cause variable
causality(var.a, cause = c("BTC")) #cause variable
```
We reject the null hypothesis and therefore show that ETH and LTC do cause BTC. 
However, BTC also causes ETH and LTC.


Forecasting with the VAR model
```{r}
fcast = predict(var.a, n.ahead = 30) # we forecast over a short horizon because beyond short horizon prediction becomes unreliable or uniform
par(mar = c(2.5,2.5,2.5,2.5))
plot(fcast)
```

Extracting the BTC forecast
```{r}
# Forecasting the DAX index
BTC.fcast = fcast$fcst[1]
head(BTC.fcast)

BTC.preds <- BTC.fcast$BTC[,1]; # extract just the predictions
BTC.preds # one month predictions
```

To get the data to the original scale we need to invert the time series
```{r}
#since the values are just difference from the previous value, to get the values on the original scale we add the last value from the BTC time series to the predicted values.
tail(cryptos)[6,1]
BTC.preds = cumsum(BTC.preds) + as.numeric(tail(cryptos)[6,1])
BTC.preds = exp(BTC.preds)
BTC.preds = ts(BTC.preds, start = c(2020, 1), frequency = 365)
plot(BTC.preds, main = "30 day BTC forecast")
```

Plot the forecast alongside the actual values
```{r}
par(mar = c(2.5,2.5,1,2.5)) #bottom, left, top, and right
plot(BTC.preds, main = "30 day BTC forecast", col = 'red', lwd = 2, ylim = c(7000, 9000))
lines(index(BTC.preds), window(exp(BTC_USD.test), end=c(2020, 30)), col = 'black', lwd = 2)
legend("topleft", legend=c("Actual", "Preds"),
       col=c("black", "red"), lty=1, cex=1)
```

Calculate errors and metrics
```{r}
accuracy(window(exp(BTC_USD.test), end=c(2020, 30)), BTC.preds)

errors <- window(BTC_USD.test, end=c(2020, 30)) - BTC.preds
checkresiduals(errors)

# sMAPE
library(TSPred)
sMAPE(window(exp(BTC_USD.test), end=c(2020, 30)), BTC.preds)

# Training residuals
checkresiduals(var.a$varresult$BTC$residuals)

```



## VARMA() method from class

Run a multivariate model using VARMA().
```{r}
m1 <- VARMA(cryptos)
summary(m1)
acf(resid(m1))
```

References: https://bookdown.org/singh_pratap_tejendra/intro_time_series_r/multivariate-ts-analysis.html












