---
title: "Untitled"
author: "Manaswi Mishra"
date: "6/1/2020"
output: html_document
---
Time Series Project - Quandl Data using Prophet

## Read Data and Import Libraries

```{r} 

library(tseries)
library(prophet)
library(forecast)
library(Quandl)
library(stringr)

```
Downloading data from Quandl
``` {r}

Quandl.api_key("UqYbx-p5DwWzyT8AH2c7")
btc_usd<-Quandl("BCHAIN/MKPRU") #Market price BTC to USD

```

Convert to chronological order and time series object for visualization
purpose. (Prophet works with dataframe and not time series objects)

Visualize basic plots for the 3 different time series objects

``` {r}
library(xts)
btc_usd_ts <- xts(btc_usd[,2], order.by=as.Date(btc_usd[,1], "%Y-%m-%d"))
autoplot(btc_usd_ts)

#Reducing the time frame as pre-2017 there was hardly any information in the time series
btc_usd_2017_2020<-btc_usd_ts["2017-01-01/"]
autoplot(btc_usd_2017_2020)
```

```{r}
#Confirming Start and End date
chron <- rev(btc_usd$Value)
c(start = tail(btc_usd$Date, 1), end = head(btc_usd$Date, 1))

```

```{r}
#Create train test for the TS

#Find row numbers of relevent dates
train_row<-which(btc_usd$Date == "2019-12-31")
test_row<-which(btc_usd$Date == "2020-01-31")



#Find start of 2016
base_date<-which(btc_usd$Date == "2016-06-01")


btc_usd_train<-btc_usd[(train_row):(base_date) , ]
btc_usd_test<-btc_usd[(test_row):(train_row-1) , ]

btc_usd_test_ts <- xts(btc_usd_test[,2], order.by=as.Date(btc_usd_test[,1], "%Y-%m-%d"))
autoplot(btc_usd_test_ts)

btc_uds_train_ts<-xts(btc_usd_train[,2], order.by=as.Date(btc_usd_train[,1], "%Y-%m-%d"))
autoplot(btc_uds_train_ts,)

```
```{r}
#Try and look into the decomposed TS
rev_ts<-btc_usd_train[nrow(btc_usd_train):1,]
rev_ts_btc = ts(rev_ts$Value ,frequency = 365)

decompose_ts = decompose(rev_ts_btc, "additive")

plot(decompose_ts)

rev_ts_btc_m<-msts(rev_ts$Value,seasonal.periods = c(7,365.25),start=c(2016,182))
decompose_ts_m = decompose(rev_ts_btc_m, "multiplicative")

plot(decompose_ts_m)
```

Try a TBATS model for basic fitting
```{r}
#fit <- tbats(rev_ts_btc_m)
#fc <- forecast(fit,h=30)
#plot(fc)
```


## Prophet Model for BTC-USD

```{r}
#Build function to customize dataframe as Prophet input, put start year for the starting year of subset
prophet_df <- function(dat,year){
  colnames(dat) <- c("ds", "y")
  print(summary(dat))
  new = subset(dat, format(as.Date(ds),"%Y")>=year)
  lam = BoxCox.lambda(new$y, method = "loglik")
  new$y = BoxCox(new$y, lam)
  fin_list = list('lam'=lam,'df'=new)
  return(fin_list)
}
btc_usd_prop = prophet_df(btc_usd_train, year = "2016")


```

``` {r}
m1 = prophet(btc_usd_prop$df, changepoint.prior.scale =0.0129, seasonality.mode = "multiplicative", changepoint.range = 0.90 )

#0.0129
#, changepoints = c('2017-01-01','2018-01-01','2019-01-01','2019-06-01','2019-12-30'),changepoint.range = 1 
#, yearly.seasonality = 5,                 
#try for 100% learn and then yearly seasonality  ,yearly.seasonality =TRUE, seasonality.mode = "additive"
# ,yearly.seasonality = TRUE
#
```



``` {r}
future <- make_future_dataframe(m1, periods = 180)
forecast.m1 <- predict(m1, future)
plot(m1,forecast.m1,xlabel = 'Time',ylabel = 'Log(BTC price)')+add_changepoints_to_plot(m1)

```




```{r}

#Inverse values using the inverse box cox

inverse_forecast <- forecast.m1
inverse_forecast$yhat_untransformed = InvBoxCox(forecast.m1$yhat, btc_usd_prop$lam)

final_forecast<-xts(inverse_forecast['yhat_untransformed'], order.by=as.Date(inverse_forecast[,1], "%Y-%m-%d"))

predict_final_forecast<-final_forecast["2020/2020-01"]



#plotting together
#Convert to Zoo so as to embed 2 plots
predict_final_forecast_zoo<-zoo(ts(predict_final_forecast,start=c(2020,1),frequency=365))


btc_usd_test_ts_zoo<-zoo(ts(btc_usd_test_ts,start=c(2020,1),frequency=365))

plot.zoo(cbind(predict_final_forecast_zoo, btc_usd_test_ts_zoo), 
         plot.type = "single", 
         col = c("red", "blue"),xlab = 'Time Jan 2020', ylab='BTC price')


```


```{r}
prophet_plot_components(m1,forecast.m1)
```


```{r}
tail(forecast.m1[,c('ds', 'yhat', 'yhat_lower', 'yhat_upper')])
```


Analyze Residual for BTC-USD model

``` {r}
#reversing the row order of the TS
a<-apply(as.matrix(btc_usd_prop$df$y),2,rev)
plot(a,type = "l")

m1.resid.new<-ts(a-forecast.m1$yhat[1:nrow(btc_usd_prop$df)],start=c(2016,182), frequency = 365.25)
ts.plot(m1.resid.new)
summary(m1.resid.new)
#checkresiduals(m1)
```

``` {r}
hist(m1.resid.new)
qqnorm(m1.resid.new, pch = 1)
qqline(m1.resid.new, col = "steelblue", lwd = 2)
```
```{r}
acf(m1.resid.new,lag=100)

```

``` {r}
adf.test(m1.resid.new)
```

``` {r}
kpss.test(m1.resid.new)
```


``` {r}
shapiro.test(m1.resid.new)
```

Residuals are stationary but the distribution is not normal (looking at
the QQ plot and Shapiro Wilk test results)


```{r}
#Creating Cross Validation metric

df.cv<-cross_validation(m1, initial = 720, period=15, horizon = 30, units = 'days')

df.p <- performance_metrics(df.cv)
df.p
plot(df.p$mae,xlab = "Days",ylab = "log(MAE) value",type="l")
```

```{r}
# Validation using MAE
performace<-df.p[c(1,4)]

invBoxCox <- function(x, lambda)
    		if (lambda == 0) exp(x) else (lambda*x + 1)^(1/lambda)

performace$mae<-invBoxCox(performace$mae,btc_usd_prop$lam)

plot(performace,type="l")

```

```{r}
#sMAPE
library(yardstick)

print("Prophet sMAPE value for the month of Jan 2020")
(sMAPE.prop<- smape_vec(as.numeric(btc_usd_test_ts_zoo),as.numeric(predict_final_forecast_zoo)))

library(regression_metrics)
(MAE.prop<- abs(mean((btc_usd_test_ts_zoo),(predict_final_forecast_zoo))))

library(forecast)
accuracy(ts(btc_usd_test_ts_zoo), ts(predict_final_forecast_zoo))

```

```{r}
test2<-btc_usd_ts["2020-01-01/2020-06-01"]
predicted_xts <- xts(forecast.m1['yhat'], order.by=as.Date(forecast.m1[,1], "%Y-%m-%d"))

predicted_xts_2020<-predicted_xts["2020-01-01/2020-06-01"]
invBoxCox <- function(x, lambda)
    		if (lambda == 0) exp(x) else (lambda*x + 1)^(1/lambda)

predicted_xts_2020$yhat<-invBoxCox(predicted_xts_2020$yhat,btc_usd_prop$lam)


#plotting together
#Convert to Zoo so as to embed 2 plots
a1<-zoo(predicted_xts_2020)


a2<-zoo(test2)

plot.zoo(cbind(a1, a2), 
         plot.type = "single", 
         col = c("red", "blue"),xlab = 'Time', ylab='BTC price/Predicted Price')
```